---
title: "Eddie Module Soil Respiration"
output: html_notebook
---


# Install the basic libraries
```{r install libraries}
# install.packages("tidyverse")
# install.packages("readxl")
# install.packages("janitor")
# install.packages("lubridate")
# install.packages("broom")
# install.packages("patchwork")
# install.packages("skimr")
```

# Load libraries    
Load the libraries each time you run a script or program   
```{r load libraries, message=FALSE, warning=FALSE}
library(tidyverse) # loads a lot of libraries in this one to add more functionality
library(readxl) # allows you to read in excel files
library(janitor) # cleans up column names and removes empty columns if wanted
library(lubridate) # allows easy conversion of varaibles to date format
library(broom) # cleans up output for easy presentation
library(patchwork)
library(skimr)
```

           
# Read in files
Decar River in Palo Iowa 15 mintue data
```{r}
soil.df <- read_excel("../data/soil_respiration_module_data.xlsx",
                    sheet = "Data for annual timeseries")
```

# Plot of soil respiration
```{r}
soil.df %>% 
  ggplot(aes(day_of_year, respiration_um_co2_m2_second)) +
  geom_line() +
  geom_point()
```

# Minimum soil respiration
```{r}
soil.df %>% 
  summarize(
    min_respiration = min(respiration_um_co2_m2_second, na.rm = TRUE),
    max_respiration = max(respiration_um_co2_m2_second, na.rm = TRUE),
    mean_respiration = mean(respiration_um_co2_m2_second, na.rm = TRUE),
    )
```



# Plot of soil temp
```{r}
soil.df %>% 
  ggplot(aes(day_of_year, soil_temp_c)) +
  geom_line() +
  geom_point()
```

# Minimum soil temp
```{r}
soil.df %>% 
  summarize(
    min_temp_c = min(soil_temp_c, na.rm = TRUE),
    max_temp_c = max(soil_temp_c, na.rm = TRUE),
    mean_temp_c = mean(soil_temp_c, na.rm = TRUE),
    )
```



# Plot of soil moisture
```{r}
soil.df %>% 
  ggplot(aes(day_of_year, soil_moisture_pct)) +
  geom_line() +
  geom_point()
```

```{r}
soil.df %>% 
  summarize(
    min_soil_moisture_pct = min(soil_moisture_pct, na.rm = TRUE),
    max_soil_moisture_pct = max(soil_moisture_pct, na.rm = TRUE),
    mean_soil_moisture_pct = mean(soil_moisture_pct, na.rm = TRUE),
    )
```




Demonstrate fitting a linear regression curve to the soil CO2 emission rate. Use % soil moisture (column D) as the independent (predictor) variable. Talk through the basics: What is a linear equation? What does the slope mean? What is R2? What is a p value? What does it mean that there is a significant negative correlation between soil moisture and respiration rate? 

# Respiration versus temperature
```{r}
soil.df %>% 
  ggplot(aes(x=soil_temp_c, y=respiration_um_co2_m2_second )) +
  geom_point() +
  geom_smooth(method="lm")
```

# Linear Model of Respiration vs Temperature
```{r}
resp_temp.model <- lm(respiration_um_co2_m2_second ~ soil_temp_c, data = soil.df)

summary(resp_temp.model)

```



```{r}
soil.df %>% 
  ggplot(aes(x=soil_moisture_pct, y=respiration_um_co2_m2_second )) +
  geom_point()
```

```{r}
soil.df %>% 
  ggplot(aes(x=soil_moisture_pct, y=respiration_um_co2_m2_second )) +
  geom_point() +
  geom_smooth(method="lm")
```

# Linear Model of Respiration vs Temperature
```{r}
resp_moist.model <- lm(respiration_um_co2_m2_second ~ soil_moisture_pct, data = soil.df)

summary(resp_moist.model)

```


# Calculate Fluxes
```{r}
soil.df <- soil.df %>% 
  mutate(micromolco2_m2_day = respiration_um_co2_m2_second * 60 * 60 * 24)
```

# Calculate Annual Flux
```{r}
umole_co2.val <- soil.df %>% 
  summarize(micromolco2_m2_day = sum(micromolco2_m2_day, na.rm = TRUE))

umole_co2.val
```

1 mole = 1000000 umole

```{r}
mole_co2_m2 <- umole_co2.val * (1/1000000)
mole_co2_m2
```


44 g CO2 per mole

```{r}
g_co2_m2 <- mole_co2_m2 * 44
g_co2_m2 
```



convert to kg
1g = 0.001 kg
```{r}
kg_co2_m2 <- g_co2_m2 * 0.001
kg_co2_m2
```

2.5 x 1011 m2 in state of michigan
flux from the entire state is 2.88 x 1012 kg for the year
1 kg = 0.001 metric ton
```{r}
co2_flux_mi <- kg_co2_m2 * 2.5 * 10^11
co2_flux_mi <- co2_flux_mi * 0.001
co2_flux_mi
```


