---
title: "Eddie Module Soil Respiration"
output: html_notebook
---


# Install the basic libraries
```{r install libraries}
# install.packages("tidyverse")
# install.packages("readxl")
# install.packages("janitor")
# install.packages("lubridate")
# install.packages("broom")
# install.packages("patchwork")
# install.packages("skimr")
```

# Load libraries    
Load the libraries each time you run a script or program   
```{r load libraries, message=FALSE, warning=FALSE}
library(tidyverse) # loads a lot of libraries in this one to add more functionality
library(readxl) # allows you to read in excel files
library(janitor) # cleans up column names and removes empty columns if wanted
library(lubridate) # allows easy conversion of varaibles to date format
library(broom) # cleans up output for easy presentation
library(patchwork)
library(skimr)
```

           
# Read in files
Decar River in Palo Iowa 15 mintue data
```{r}
highres.df <- read_excel("../data/soil_respiration_module_data.xlsx",
                    sheet = "Growing season")
```

# Plot of soil respiration
```{r}
highres.df %>% 
  ggplot(aes(decimal_day, soil_respiration)) +
  geom_line() +
  geom_point()+
  scale_x_continuous(
    limits = c(200,211),
    breaks = seq(200, 211, by = 1)
  )
```



# Plot of soil temp
```{r}
highres.df %>% 
  ggplot(aes(decimal_day, soil_temp_7_5_cm_c)) +
  geom_line() +
  geom_point()+
  scale_x_continuous(
    limits = c(200,211),
    breaks = seq(200, 211, by = 1)
  )
```
# Plot of soil moisture
```{r}
highres.df %>% 
  ggplot(aes(decimal_day, soil_moisture_pct)) +
  geom_line() +
  geom_point()+
  scale_x_continuous(
    limits = c(200,211),
    breaks = seq(200, 211, by = 1)
  )
```
# Plot of air temp
```{r}
highres.df %>% 
  ggplot(aes(decimal_day, shaded_air_temp_c_2m)) +
  geom_line() +
  geom_point()+
  scale_x_continuous(
    limits = c(200,211),
    breaks = seq(200, 211, by = 1)
  )
```
# Plot of soil heat flux
```{r}
highres.df %>% 
  ggplot(aes(decimal_day, soil_heat_flux_wm2)) +
  geom_line() +
  geom_point()+
  scale_x_continuous(
    limits = c(200,211),
    breaks = seq(200, 211, by = 1)
  )
```

# Plot of rain
```{r}
highres.df %>% 
  ggplot(aes(decimal_day, hourly_rainfall_mm)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(
    limits = c(200,211),
    breaks = seq(200, 211, by = 1)
  )
```



Demonstrate fitting a linear regression curve to the soil CO2 emission rate. Use % soil moisture (column D) as the independent (predictor) variable. Talk through the basics: What is a linear equation? What does the slope mean? What is R2? What is a p value? What does it mean that there is a significant negative correlation between soil moisture and respiration rate? 

```{r}
highres.df %>% 
  ggplot(aes(x=soil_moisture_pct, y=soil_respiration )) +
  geom_point()
```

```{r}
highres.df %>% 
  ggplot(aes(x=soil_moisture_pct, y=soil_respiration )) +
  geom_point() +
  geom_smooth(method="lm")
```








